---
title: 'Recreational Survey -- October 2023'
subtitle: "October 10-11, 2023   [*F/V Fin Chaser*](https://finchasersportfishing.com/)   Ocean City, MD"
format: html
editor: visual
echo: false
message: false
warning: false
params:
  seed: 20230906
  vessel_sp_kts: 12
  start:
    value: 
      d1: '2023-10-11 06:00'
  buoys: FALSE
---

```{r}
#| label: packages

library(sf); library(ggplot2); library(concaveman); library(TSP); library(units)
library(data.table); library(targets)
```

```{r}
#| label: towers

turbines <- usw_locations |> 
  tar_read() |> 
  st_read(layer = 'WTG_Layout_2022_0111', quiet = TRUE)

turbine_boundary <- concaveman::concaveman(turbines)

setDT(turbines)[c(st_contains(turbine_boundary,
                           st_buffer(geom, units::as_units(10, 'meters')),
                           sparse = F)), location := 'inner']
turbines[is.na(location), location := 'outer']

zones <- st_read(tar_read(usw_buildout),
                 quiet = T)
zones <- zones[zones$zone == 'A', ]
```

```{r}
#| label: stations
set.seed(params$seed)

exp_sites <- turbines |>
  DT(grepl('^[JKLMNOPQR]', ID) & location == 'inner') |>
  # first selected will be surveyed the first day,
  #   second selected will be surveyed the second day
  DT(sample(.N, 1)) |>
  DT(, .(station = ID, geometry = geom)) |>
  st_as_sf()

ctrl_sites <- data.table(
  station = c('Twin Wrecks', 'Washingtonian'),
  lat = c(38.23071, 38.43049),
  long = c(-74.75808, -74.76770)
)

ctrl_sites <- 
  # Randomize order by day
  ctrl_sites[sample(1:nrow(ctrl_sites))] |> 
  st_as_sf(coords = c('long', 'lat'),
           crs = 4326)

stations <- rbind(ctrl_sites, exp_sites)

ocmd <- c(-75.10333, 38.32742) %>% 
  st_point() %>% 
  st_sfc(crs = 4326) %>% 
  st_sf(geometry = .,
        station = 'ocmd')

stations <- rbind(ocmd, stations)

if(params$buoys == TRUE){
  buoys <- data.table(
    station = c('rtwb', 'metocean'),
    lat = c(38.303, 38.35277),
    long = c(-74.645, -74.75365)
  ) |> 
    st_as_sf(coords = c('long', 'lat'),
             crs = 4326)
  
  stations <- rbind(stations, buoys)
}

setDT(stations)[, ':='(lon = st_coordinates(geometry)[,1],
                       lat = st_coordinates(geometry)[,2])]
dm <- function(x){
  paste(as.integer(x), round((abs(x) %% 1) * 60, 4))
}
stations[, ':='(lon_dm = dm(lon),
                lat_dm = dm(lat))]

stations <- st_as_sf(stations)
```

```{r}
tsp <- TSP(as.dist(st_distance(stations)))

tour <- solve_TSP(tsp, method = 'farthest_insertion')

path <- c(cut_tour(tour, 1, exclude_cut = FALSE), 1)

trip <- list(stations)
```

```{r}
#| label: calc-dist-matrices

dist_calc <- function(stations, new_unit = 'nautical_mile'){
  dist_mat <- st_distance(stations)
  
  dist_mat <- set_units(dist_mat, new_unit, mode = 'standard')
  
  rownames(dist_mat) <- stations$station
  colnames(dist_mat) <- stations$station

  dist_mat
}

tf <- function(dist_mat, speed, pretty = T){
  mat <- dist_mat |>
    as.numeric() |>
    matrix(ncol = sqrt(length(dist_mat)))
  
  if(pretty == T){
  mat <- (mat / speed) * 3600
  
  mat <- ceiling(mat / (5 * 60)) * (5 * 60)
  
  mat <- as.POSIXct(mat,
                    tz = 'UTC',
                    origin = '1970-01-01')
  
  mat <- format(mat, '%H h %M m')
  
  mat <- matrix(mat, ncol = sqrt(length(mat)))
  
  diag(mat) <- '-'
  
  }else{
    mat <- ceiling((mat / speed) * 12) / 12
    
    mat <- matrix(mat, ncol = sqrt(length(mat)))
  }
  
  rownames(mat) <- rownames(dist_mat)
  colnames(mat) <- colnames(dist_mat)
  
  mat
}

nm_matrix <- round(dist_calc(stations), 2)
time_matrix_raw <- tf(nm_matrix, params$vessel_sp_kts, pretty = F)
time_matrix_pretty <- tf(nm_matrix, params$vessel_sp_kts)
```

```{r}
#| label: cruise-seq
survey_seq <- list()

for(i in 1){
  n_stations <- length(trip[[i]]$station)
  
  survey_seq[[i]] <-
    data.table(start = stations$station[path[1:(length(path) - 1)]],
                         end = stations$station[path[2:length(path)]],
               sttime = c(as.POSIXct(params$start[[i]]),
                          rep(NA, n_stations - 1)),
               endtime = c(as.POSIXct(params$start[[i]]),
                           rep(NA, n_stations - 1)),
               jig = c(as.POSIXct(params$start[[i]]),
                       rep(NA, n_stations - 1)),
               drop = c(as.POSIXct(params$start[[i]]),
                        rep(NA, n_stations - 1)))
  
  # travel time
  survey_seq[[i]][1, endtime := sttime + time_matrix_raw[start, end] * 3600]
  if(survey_seq[[i]][1, end] %in% c('rtwb', 'metocean')){
    # no jig time
    survey_seq[[i]][1, jig := endtime]
    # drop time stands in for 10 min of videos/pictures
    survey_seq[[i]][1, drop := jig + 10 * 60]
  }else{
    # 5 min prep plus jig time (15 min) plus 5 min changeover
    survey_seq[[i]][1, jig := endtime + (5 + 15 + 5) * 60]
    # drop time (3 min) plus 2 min changeover
    survey_seq[[i]][1, drop := jig + 15 * (3 + 2) * 60]
  }

  
  for(j in 2:n_stations){
    survey_seq[[i]][j, sttime := survey_seq[[i]]$drop[j - 1]]
    # travel time
    survey_seq[[i]][j, endtime := sttime + time_matrix_raw[start, end] * 3600]
    if(survey_seq[[i]][j, end] %in% c('rtwb', 'metocean')){
      # no jig time
      survey_seq[[i]][j, jig := endtime]
      # drop time stands in for 10 min of videos/pictures
      survey_seq[[i]][j, drop := jig + 10 * 60]
    }else{
      # jig time (15 min) plus 5 min changeover
      survey_seq[[i]][j, jig := endtime + (5 + 15 + 5) * 60]
      # drop time (3 min) plus 2 min changeover
      survey_seq[[i]][j, drop := jig + 15 * (3 + 2) * 60]
    }
  }
}
```

## Participants

::: columns
::: {.column width="50%"}
### `r format(as.POSIXct(params$start$d1) - 24*60*60, '%d %B')`

-   [Evan Kostelecky](mailto:ekostelecky@umces.edu)
-   [Kirsten Silva](mailto:ksilva@umces.edu)
:::

::: {.column width="50%"}
### `r format(as.POSIXct(params$start$d1), '%d %B')`

-   [Evan Kostelecky](mailto:ekostelecky@umces.edu)
-   [Kirsten Silva](mailto:ksilva@umces.edu)
-   [Tom Cole](mailto:tcole007@yahoo.com)
-   [Capt. Dan Stauffer](mailto:finchasersportfishing@gmail.com)
-   Dave Weller, *Fin Chaser* mate
:::
:::

## Schedule

### `r format(as.POSIXct(params$start$d1) - 24*60*60, '%d %B')`

-   Scientific crew (EK, KS)
    -   Pack CBL truck (2017 Silverado) at Solomons, MD. Drive to West Ocean City, MD
    -   Purchase dry ice at Safeway in Chester, MD
        -   1925 Main St, Chester, MD 21619
    -   Check into the Microtel in West Ocean City, MD

### `r format(as.POSIXct(params$start$d1), '%d %B')`

-   Full crew (EK. KS, TC, DS, DW mate)
    -   `r paste(format(survey_seq[[1]]$sttime[1], '%H:%M'), format(survey_seq[[1]]$endtime[nrow(survey_seq[[1]])], '%H:%M'), sep = ' - ')`: Recreational fishing survey aboard F/V Fin Chaser
        -   312 Talbot St, Ocean City, MD 21842
-   Scientific crew
    -   Return to Solomons, MD.
    -   `r format(survey_seq[[1]]$endtime[nrow(survey_seq[[1]])] + 4.5*60*60, '%H:%M')`: Arrive in Solomons. Unpack and return truck.

## Packing list

::: columns
::: {.column width="33%"}
🔲 3 fishing rods

🔲 Large cooler

🔲 Sampling bags and bag tags

🔲 Data sheets/clipboard
:::

::: {.column width="33%"}
🔲 3 buckets

🔲 GPS

🔲 Stopwatch

🔲 Castaway CTD
:::

::: {.column width="33%"}
🔲 MS222

🔲 Measuring board

🔲 Weight bucket with scales
:::
:::

## Protocols

::: columns
::: {.column width="50%"}
### GPS

-   Set waypoint prefix to "YYMMRec"
    -   From the main menu, select Setup, then Waypoints, then change prefix
-   To start recording track, from map screen, press **ENTER** twice
-   Mark waypoints when jigging or drop starts or ends
    -   Press **MARK** then **ENTER**
-   To stop recording track
    -   From map screen, press **ENTER**, select **STOP**, then **SAVE**

### Data timing

-   Data Recorder (DR) confirms that Captain is ready to begin;
-   starts recording route;
-   tells fishers to start and simultaneously marks GPS waypoint;
-   starts timer;
-   marks start waypoint and time on data sheet;
-   says "STOP" when time is up and simultaneously marks GPS waypoint;
-   stops recording route;
-   stops timer;
-   marks end waypoint on data sheet.
:::

::: {.column width="50%"}
### Data recording

-   No-catch completed jigs/drops are assumed.
-   Record if jig/drop is not completed or if a fish is captured
-   Repeated lines for a jig/drop are only necessary if more than one angler lands a fish or does not complete the jig/drop.

### Jig

-   Jig survey is conducted by the Mate for 15 minutes.
-   Jigging is resumed if a fish is landed until 15 minutes of active fishing are complete.
-   Time is paused while jig is out of the water.
-   All fish are weighed and measured.
-   First 10 black sea bass are sacrificed.

### Drop

-   15 x 3 minute drops each by 3 anglers, one of which must be the Mate
-   All fish are weighed and measured.
-   First 10 black sea bass are sacrificed.
-   Fishing DOES NOT resume if fish(es) are landed.
:::
:::

\newpage

```{r}
library(leaflet)
wms_enc <- 'https://gis.charttools.noaa.gov/arcgis/rest/services/MCS/ENCOnline/MapServer/exts/MaritimeChartService/WMSServer'
leaflet() |> 
  setView(lng = -74.9, lat = 38.25, zoom = 10) |> 
  addWMSTiles(
    wms_enc,
    layers = c(0:3, 6),
    options = WMSTileOptions(format = "image/png", transparent = T)
  ) |> 
  addPolygons(data = zones, color = 'pink') |>
  addCircleMarkers(data = stations, label = ~ station) |> 
  addPolylines(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(trip[[1]])[path]),
      'LINESTRING'
    ), crs = 4326
  ), color = 'blue') |>  
  addMeasure(primaryLengthUnit = 'meters', primaryAreaUnit = 'hectares') |>
  addMiniMap(zoomLevelFixed = 7)
```

## Cruise plan

Times calculated at `r params$vessel_sp_kts` knots.

### `r format(as.POSIXct(params$start$d1), '%d %B')`

::: columns
::: {.column width="50%"}
-   `r format(survey_seq[[1]]$sttime[1] - 30 * 60, '%H:%M')`: DHS, KS, DW meet at Talbot St Pier, load gear onto *F/V Fin Chaser*
-   `r format(survey_seq[[1]]$sttime[1], '%H:%M')`: *Fin Chaser* depart Talbot St Pier, transit to **`r survey_seq[[1]]$end[1]`**
    -   <font size= "5"> **`r stations[stations$station == survey_seq[[1]]$end[1],]$lat_dm`, `r stations[stations$station == survey_seq[[1]]$end[1],]$lon_dm`** </font>
-   `r format(survey_seq[[1]]$endtime[1], '%H:%M')`: Arrive `r survey_seq[[1]]$end[1]`.
    -   Mate arranges gear for jigging.
    -   Sci crew organizes CTD and MS-222 bucket
    -   Environmental data recorded
-   `r format(survey_seq[[1]]$endtime[1] + 5*60, '%H:%M')`: Mate begins jigging. Scientific crew deploys CTD.
-   `r format(survey_seq[[1]]$jig[1] - 5*60, '%H:%M')` Mate ends jigging.
-   `r format(survey_seq[[1]]$jig[1], '%H:%M')`: Survey crew begins drop survey.
:::

::: {.column width="50%"}
-   `r format(survey_seq[[1]]$drop[1], '%H:%M')`: Drop survey ends. Transit to **`r survey_seq[[1]]$end[2]`**.
    -   <font size= "5"> **`r stations[stations$station == survey_seq[[1]]$end[2],]$lat_dm`, `r stations[stations$station == survey_seq[[1]]$end[2],]$lon_dm`** </font>
-   `r format(survey_seq[[1]]$endtime[2], '%H:%M')`: Arrive `r survey_seq[[1]]$end[2]`. Prep as before.
-   `r format(survey_seq[[1]]$endtime[2] + 5*60, '%H:%M')`: Mate begins jigging.
-   `r format(survey_seq[[1]]$jig[2] - 5*60, '%H:%M')`: Mate ends jigging.
-   `r format(survey_seq[[1]]$jig[2], '%H:%M')`: Survey crew begins drop survey
-   `r format(survey_seq[[1]]$drop[2], '%H:%M')`: Drop survey ends. Transit to **`r survey_seq[[1]]$end[3]`**.
    -   <font size= "5"> **`r stations[stations$station == survey_seq[[1]]$end[3],]$lat_dm`, `r stations[stations$station == survey_seq[[1]]$end[3],]$lon_dm`** </font>
-   `r format(survey_seq[[1]]$endtime[3], '%H:%M')`: Arrive `r survey_seq[[1]]$end[3]`. Prep as before.
-   `r format(survey_seq[[1]]$endtime[3] + 5*60, '%H:%M')`: Mate begins jigging.
-   `r format(survey_seq[[1]]$jig[3] - 5*60, '%H:%M')`: Mate ends jigging.
-   `r format(survey_seq[[1]]$jig[3], '%H:%M')`: Survey crew begins drop survey
-   `r format(survey_seq[[1]]$drop[3], '%H:%M')`: Drop survey ends. Return to Talbot St Pier.
-   `r format(survey_seq[[1]]$endtime[4], '%H:%M')`: Arrive Talbot St Pier. Unload gear.
:::
:::

## Distances

```{r}
#| label: dist-matrix
nm_matrix |>
  as.numeric() |>
  matrix(ncol = length(stations$station),
         dimnames = list(stations$station, stations$station)) |>
  knitr::kable('html', align = 'c', caption = 'Distance, nautical miles')

time_matrix_pretty |>
  matrix(ncol = length(stations$station),
         dimnames = list(stations$station, stations$station)) |> 
  knitr::kable('html', align = 'c', caption = paste('Distance, time at', 
                                                    params$vessel_sp_kts,
                                                    'kts')
  )
```
