---
title: 'Pot Survey -- May 2023'
subtitle: "May 1 - 3, 2023   *F/V Integrity*   West Ocean City, MD"
format: html
editor: visual
echo: false
message: false
warning: false
params:
  n_control: 2
  n_exp: 4
  seed: 20230501
  vessel_sp_kts: 8
  d1_start: '2023-05-02 02:30'
  d2_start: '2023-05-03 02:30'
---

```{r}
#| label: params
library(sf); library(ggplot2); library(concaveman); library(TSP); library(units)
library(data.table); library(targets)
```

```{r}
#| label: geo-base

turbines <- usw_locations |> 
  tar_read() |> 
  st_read(layer = 'WTG_Layout_2022_0111', quiet = TRUE)

turbine_boundary <- concaveman::concaveman(turbines)

setDT(turbines)[c(st_contains(turbine_boundary,
                           st_buffer(geom, units::as_units(10, 'meters')),
                           sparse = F)), location := 'inner']
turbines[is.na(location), location := 'outer']


zones <- st_read(tar_read(usw_buildout),
                 quiet = T)
zones <- zones[zones$zone == 'A', ]
ctrl_area <- st_read(tar_read(frm_control),
                     quiet = T)
```

```{r}
#| label: ctrl-site-selection
set.seed(params$seed)

ctrl_sites <- ctrl_area |> 
  st_transform(26918) |>
  st_buffer(dist = -units::set_units(550, 'meter')) |>
  st_sample(params$n_control)|> 
  st_as_sf() |> 
  st_transform(4326) |>
  DT(order(st_coordinates(x)[, 1])) |> 
  DT(, station := paste0('C', substr(params$seed, 3, 6),
                         '-',
                         1:params$n_control)) 

setnames(ctrl_sites, 'x', 'geometry')
st_geometry(ctrl_sites) <- 'geometry'
```

```{r}
#| label: exp-site-selection

exp_sites <- turbines |>
  DT(grepl('^[JKLMNOPQR]', ID) & location == 'inner') |>
  DT(sample(.N, params$n_exp)) |>
  DT(, .(station = ID, geometry = geom)) |>
  st_as_sf()

```

```{r}
#| label: combine-sites
ocmd <- c(-75.10333, 38.32742) %>% 
  st_point() %>% 
  st_sfc(crs = 4326) %>% 
  st_sf(geometry = .,
        station = 'OCMD')

stations <- rbind(ocmd, ctrl_sites, exp_sites)

setDT(stations)[, ':='(lon = st_coordinates(geometry)[,1],
                       lat = st_coordinates(geometry)[,2])]
dm <- function(x){
  paste(as.integer(x), round((abs(x) %% 1) * 60, 4))
}
stations[, ':='(lon_dm = dm(lon),
                lat_dm = dm(lat))]

stations <- st_as_sf(stations)
```

```{r}
#| label: tsp

tsp <- TSP(as.dist(st_distance(stations)))

tour <- solve_TSP(tsp, method = 'farthest_insertion')

path <- c(cut_tour(tour, 1, exclude_cut = FALSE), 1)
```

```{r}
#| label: calc-dist-matrices
dist_calc <- function(stations, new_unit = 'nautical_mile'){
  dist_mat <- st_distance(stations)
  
  dist_mat <- set_units(dist_mat, new_unit, mode = 'standard')
  
  rownames(dist_mat) <- stations$station
  colnames(dist_mat) <- stations$station

  dist_mat
}

tf <- function(dist_mat, speed, pretty = T){
  mat <- dist_mat |>
    as.numeric() |>
    matrix(ncol = sqrt(length(dist_mat)))
  
  if(pretty == T){
  mat <- (mat / speed) * 3600
  
  mat <- ceiling(mat / (5 * 60)) * (5 * 60)
  
  mat <- as.POSIXct(mat,
                    tz = 'UTC',
                    origin = '1970-01-01')
  
  mat <- format(mat, '%H h %M m')
  
  mat <- matrix(mat, ncol = sqrt(length(mat)))
  
  diag(mat) <- '-'
  
  }else{
    mat <- ceiling((mat / speed) * 12) / 12
    
    mat <- matrix(mat, ncol = sqrt(length(mat)))
  }
  
  rownames(mat) <- rownames(dist_mat)
  colnames(mat) <- colnames(dist_mat)
  
  mat
}

nm_matrix <- round(dist_calc(stations), 2)
time_matrix_raw <- tf(nm_matrix, params$vessel_sp_kts, pretty = F)
time_matrix_pretty <- tf(nm_matrix, params$vessel_sp_kts)
```

```{r}
#| label: cruise-seq

deploy_seq <- data.table(start = stations$station[path[1:(length(path) - 1)]],
                         end = stations$station[path[2:length(path)]],
                         sttime = c(as.POSIXct(params$d1_start) + 30*60, rep(NA, 6)),
                         endtime = c(as.POSIXct(params$d1_start), rep(NA, 6)),
                         pots = c(as.POSIXct(params$d1_start), rep(NA, 6)))
deploy_seq[1, endtime := sttime + time_matrix_raw[start, end] * 3600]
deploy_seq[1, pots := endtime + 10 * 60]

for(i in 2:(length(path)-1)){
  deploy_seq[i, sttime := deploy_seq$pots[i-1]]
  deploy_seq[i, endtime := sttime + time_matrix_raw[start, end] * 3600]
  deploy_seq[i, pots := endtime + 10 * 60]
}



rec_seq <- data.table(start = stations$station[path[1:(length(path) - 1)]],
                         end = stations$station[path[2:length(path)]],
                         sttime = c(as.POSIXct(params$d2_start) + 30*60, rep(NA, 6)),
                         endtime = c(as.POSIXct(params$d2_start), rep(NA, 6)),
                         pots = c(as.POSIXct(params$d2_start), rep(NA, 6)))
rec_seq[1, endtime := sttime + time_matrix_raw[start, end] * 3600]
rec_seq[1, pots := endtime + 15 * 60]

for(i in 2:(length(path)-1)){
  rec_seq[i, sttime := rec_seq$pots[i-1]]
  rec_seq[i, endtime := sttime + time_matrix_raw[start, end] * 3600]
  rec_seq[i, pots := endtime + 15 * 60]
}
```

```{r}
library(leaflet)
wms_enc <- 'https://gis.charttools.noaa.gov/arcgis/rest/services/MCS/ENCOnline/MapServer/exts/MaritimeChartService/WMSServer'
leaflet() |> 
  setView(lng = -74.9, lat = 38.25, zoom = 10) |> 
  addWMSTiles(
    wms_enc,
    layers = c(0:3, 6),
    options = WMSTileOptions(format = "image/png", transparent = T)
  ) |> 
  addPolygons(data = zones, color = 'pink') |>
  addPolygons(data = ctrl_area, color = 'gray')|>
  addCircleMarkers(data = stations, label = ~ station) |> 
  addPolylines(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[c(path),])),
      'LINESTRING'
    ), crs = 4326
  )) |>
  addMeasure(primaryLengthUnit = 'meters', primaryAreaUnit = 'hectares') |>
  addMiniMap(zoomLevelFixed = 7)
```

## Packing list

::: columns
::: {.column width="33%"}
🔲 Large cooler

🔲 Sampling bags and bag tags

🔲 Data sheets/clipboard

🔲 \>2 fish pots
:::

::: {.column width="33%"}
🔲 Two kill buckets

🔲 GPS

🔲 HOBO loggers (2)

🔲 MS222
:::

::: {.column width="33%"}
🔲 Measuring board

🔲 Weight bucket with scales

🔲 2 fish totes
:::
:::

## Participants

::: columns
::: {.column width="50%"}
### May 1-3

-   [Mike O'Brien](mailto:%20obrien@umces.edu): 267-970-1973
-   [Lauren Rodriguez](mailto:%20lrodriguez@umces.edu): 586-904-1264
:::

::: {.column width="50%"}
### May 2-3

-   [Mike O'Brien](mailto:%20obrien@umces.edu)
-   [Lauren Rodriguez](mailto:%20lrodriguez@umces.edu)
-   Capt. Brandon Harrington
-   *Integrity* mate
:::
:::

## Schedule

### 1 May

Scientific crew (MOB, LR)

-   0800 - 0900 MOB pack CBL truck (2017 Silverado) at Solomons, MD. Drive to Harry S Truman Park and Ride.
    -   200 Harry S. Truman Pkwy, Annapolis, MD 21401
-   1015 - 1230 MOB picks up LR. Drive to MD Coastal Bays Program office.
    -   8219 Stephen Decatur Hwy, Berlin, MD 21811
-   1230 - 1250 Load pots and two rope drums onto trailer.
-   1250 - 1300 Drive to Sea Born LLC.
    -   12906 Sunset Ave, Ocean City, MD 21842

Full crew (MOB, LR, BH, mate)

-   1300 - 1330 Unload pots. KH/mate begin loading pots onto vessel.
-   1330 - 1530 DHS/MOB drive to MD Coastal Bays Program. Repeat pot loading/return/unloading.
-   1530 - 1600 Drop trailer off at MDCBP. Return to W Ocean City.
-   1600 - 1800 Finish rigging ground line.
-   1800 Check into Microtel
    -   12536 Ocean Gateway, Ocean City, MD 21842

### `r format(as.POSIXct(params$d1_start), '%B %d')`

Scientific & Sea Born crew

-   `r format(as.POSIXct(params$d1_start), '%H:%M')` Meet at *F/V Integrity*
-   `r format(deploy_seq$sttime[1], '%H:%M')` Depart `r deploy_seq$start[1]` for **`r deploy_seq$end[1]`**
-   <font size= "5"> **`r stations[stations$station == deploy_seq$end[1],]$lat_dm`, `r stations[stations$station == deploy_seq$end[1],]$lon_dm`** </font>
-   `r format(deploy_seq$endtime[1], '%H:%M')` Arrive **`r deploy_seq$end[1]`**. Deploy pots.
-   `r format(deploy_seq$pots[1], '%H:%M')` Finish deploying pots. Depart for **`r deploy_seq$end[2]`**
-   <font size= "5"> **`r stations[stations$station == deploy_seq$end[2],]$lat_dm`, `r stations[stations$station == deploy_seq$end[2],]$lon_dm`** </font>
-   `r format(deploy_seq$endtime[2], '%H:%M')` Arrive **`r deploy_seq$end[2]`**. Deploy pots.
-   `r format(deploy_seq$pots[2], '%H:%M')` Finish deploying pots. Depart for **`r deploy_seq$end[3]`**
-   <font size= "5"> **`r stations[stations$station == deploy_seq$end[3],]$lat_dm`, `r stations[stations$station == deploy_seq$end[3],]$lon_dm`** </font>
-   `r format(deploy_seq$endtime[3], '%H:%M')` Arrive **`r deploy_seq$end[3]`**. Deploy pots.
-   `r format(deploy_seq$pots[3], '%H:%M')` Finish deploying pots. Depart for **`r deploy_seq$end[4]`**
-   <font size= "5"> **`r stations[stations$station == deploy_seq$end[4],]$lat_dm`, `r stations[stations$station == deploy_seq$end[4],]$lon_dm`** </font>
-   `r format(deploy_seq$endtime[4], '%H:%M')` Arrive **`r deploy_seq$end[4]`**. Deploy pots.
-   `r format(deploy_seq$pots[4], '%H:%M')` Finish deploying pots. Depart for **`r deploy_seq$end[5]`**
-   <font size= "5"> **`r stations[stations$station == deploy_seq$end[5],]$lat_dm`, `r stations[stations$station == deploy_seq$end[5],]$lon_dm`** </font>
-   `r format(deploy_seq$endtime[5], '%H:%M')` Arrive **`r deploy_seq$end[5]`**. Deploy pots.
-   `r format(deploy_seq$pots[5], '%H:%M')` Finish deploying pots. Depart for **`r deploy_seq$end[6]`**
-   <font size= "5"> **`r stations[stations$station == deploy_seq$end[6],]$lat_dm`, `r stations[stations$station == deploy_seq$end[6],]$lon_dm`** </font>
-   `r format(deploy_seq$endtime[6], '%H:%M')` Arrive **`r deploy_seq$end[6]`**. Deploy pots.
-   `r format(deploy_seq$pots[6], '%H:%M')` Finish deploying pots. Depart for **`r deploy_seq$end[7]`**
-   `r format(deploy_seq$endtime[7], '%H:%M')` Arrive **`r deploy_seq$end[7]`**

Scientific crew picks up dry ice from Harris Teeter

-   31221 Americana Pkwy, Selbyville, DE 19975

### `r format(as.POSIXct(params$d2_start), '%B %d')`

Scientific & Sea Born crew

-   `r format(as.POSIXct(params$d2_start), '%H:%M')` Meet at *F/V Integrity*
-   `r format(rec_seq$sttime[1], '%H:%M')` Depart `r rec_seq$start[1]` for **`r rec_seq$end[1]`**
-   `r format(rec_seq$endtime[1], '%H:%M')` Arrive **`r rec_seq$end[1]`**. Deploy pots.
-   `r format(rec_seq$pots[1], '%H:%M')` Finish deploying pots. Depart for **`r rec_seq$end[2]`**
-   `r format(rec_seq$endtime[2], '%H:%M')` Arrive **`r rec_seq$end[2]`**. Deploy pots.
-   `r format(rec_seq$pots[2], '%H:%M')` Finish deploying pots. Depart for **`r rec_seq$end[3]`**
-   `r format(rec_seq$endtime[3], '%H:%M')` Arrive **`r rec_seq$end[3]`**. Deploy pots.
-   `r format(rec_seq$pots[3], '%H:%M')` Finish deploying pots. Depart for **`r rec_seq$end[4]`**
-   `r format(rec_seq$endtime[4], '%H:%M')` Arrive **`r rec_seq$end[4]`**. Deploy pots.
-   `r format(rec_seq$pots[4], '%H:%M')` Finish deploying pots. Depart for **`r rec_seq$end[5]`**
-   `r format(rec_seq$endtime[5], '%H:%M')` Arrive **`r rec_seq$end[5]`**. Deploy pots.
-   `r format(rec_seq$pots[5], '%H:%M')` Finish deploying pots. Depart for **`r rec_seq$end[6]`**
-   `r format(rec_seq$endtime[6], '%H:%M')` Arrive **`r rec_seq$end[6]`**. Deploy pots.
-   `r format(rec_seq$pots[6], '%H:%M')` Finish deploying pots. Depart for **`r rec_seq$end[7]`**
-   `r format(rec_seq$endtime[7], '%H:%M')` Arrive **`r rec_seq$end[7]`**
-   `r format(rec_seq$endtime[7] + 20*60, '%H:%M')` Pick up trailer from MD Coastal Bays Program
-   `r format(rec_seq$endtime[7] + 35*60, '%H:%M')` Remove pots from F/V Integrity, load onto trailer. Unload at MD Coastal Bays program. Repeat as needed.

`r format(rec_seq$endtime[7] + 2*3600, '%H:%M')` Scientific crew drive back to Solomons, MD

## Distances

```{r}
#| label: dist-matrix
nm_matrix |>
  as.numeric() |>
  matrix(ncol = length(stations$station),
         dimnames = list(stations$station, stations$station)) |>
  knitr::kable('html', align = 'c', caption = 'Distance, nautical miles')

time_matrix_pretty |>
  matrix(ncol = length(stations$station),
         dimnames = list(stations$station, stations$station)) |> 
  knitr::kable('html', align = 'c', caption = paste('Distance, time at', 
                                                    params$vessel_sp_kts,
                                                    'kts')
  )
```

```{r}
#| echo: false

library(ggplot2); library(sf); library(geoarrow)

bathy <- read_sf('data/geo/bathymetry_midatl.gpkg') |>
  st_transform(4326)

midatl <- read_geoparquet_sf('data/geo/matl_states.parquet')

ggplot() +
  geom_sf(data = bathy[bathy$elev_m < 100,]) +
  geom_sf(data = midatl) +
  geom_sf(data = stations, size = 6, fill = 'red') +
  geom_sf_label(data = stations, aes(label = station), nudge_x = 0.03, nudge_y = 0.01) +
  geom_sf(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[c(path),])),
      'LINESTRING'
      ), crs = 4326
    )) +
  labs(x = NULL, y = NULL) +
  coord_sf(xlim  = c(-75.2, -74.55),
           ylim = c(38.1, 38.5), expand = F) +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 45))

```
