---
title: 'Recreational Survey -- Shakedown August 2022'
subtitle: "August 15 - 16, 2022   [*F/V Fin Chaser*](https://finchasersportfishing.com/)   Ocean City, MD"
format: html
editor: visual
echo: false
message: false
warning: false
---

```{r}
#| label: setup
#| include: false

knitr::opts_knit$set(root.dir = 'c:/users/darpa2/analysis/frm-survey')
```

```{r}
#| label: params

library(sf); library(ggplot2); library(concaveman); library(TSP); library(units)
library(data.table); library(targets)

n_exp <- 1
chosen_ctrl <- 'Great Eastern Reef'
vessel_speed_kts <- 12

seed <- 20220815
```

```{r}
#| label: towers

towers <- usw_locations |> 
  tar_read() |> 
  st_read(layer = 'WTG_Layout_2022_0111', quiet = TRUE)

tower_boundary <- concaveman::concaveman(towers)

setDT(towers)[c(st_contains(tower_boundary,
                           st_buffer(geom, units::as_units(10, 'meters')),
                           sparse = F)), location := 'inner']
towers[is.na(location), location := 'outer']
```

```{r}
#| label: stations
set.seed(seed)

# Randomization procedure not conducted due to weather. Selecting closer site:
#   K07

# exp_sites <- towers |> 
#   DT(grepl('^[JKLMNOPQR]', ID) & location == 'inner') |>
#   DT(sample(.N, n_exp)) |> 
#   DT(, .(station = ID, geometry = geom)) |> 
#   st_as_sf()

exp_sites <- towers |> 
  DT(ID == 'K07', .(station = ID, geometry = geom)) |> 
  st_as_sf()


ctrl_sites <- data.frame(
  station = c('Twin Wrecks', 'Great Eastern Reef'),
  lat = c(38.23071, 38.43049),
  long = c(-74.75808, -74.76770)
) |> 
  st_as_sf(coords = c('long', 'lat'),
           crs = 4326) |> 
  DT(station == chosen_ctrl)

ocmd <- c(-75.10333, 38.32742) %>% 
  st_point() %>% 
  st_sfc(crs = 4326) %>% 
  st_sf(geometry = .,
        station = 'ocmd')

stations <- rbind(ocmd, ctrl_sites, exp_sites)
```

```{r}
#| label: tsp

tsp <- TSP(as.dist(st_distance(stations)))

tour <- solve_TSP(tsp, method = 'farthest_insertion')

path <- c(cut_tour(tour, 1, exclude_cut = FALSE), 1)
```

```{r}
#| label: calc-dist-matrices

tf <- function(dist_mat, speed){
  mat <- dist_mat |> 
    as.numeric() |> 
    matrix(ncol = sqrt(length(dist_mat)))
  
  mat <- (mat / speed) * 3600
  
  mat <- ceiling(mat / (5 * 60)) * (5 * 60)
  
  mat <- as.POSIXct(mat,
                    tz = 'UTC',
                    origin = '1970-01-01')
  
  mat <- format(mat, '%H h %M m')
  
  mat <- matrix(mat, ncol = sqrt(length(mat)))
  
  mat
}

nm_matrix <- round(set_units(st_distance(st_as_sf(stations)), 'nautical_mile'), 2)
time_matrix <- tf(nm_matrix, vessel_speed_kts)
```

## Participants

::: columns
::: {.column width="50%"}
### August 15

-   [Dave Secor](mailto:%20secor@umces.edu): 301-481-5053
-   [Mike O'Brien](mailto:%20obrien@umces.edu): 267-970-1973
:::

::: {.column width="50%"}
### August 16

-   [Dave Secor](mailto:%20secor@umces.edu)
-   [Mike O'Brien](mailto:%20obrien@umces.edu)
-   [Todd Sumner](mailto:%20t.sumner@uswindinc.com): 443-240-2824
-   [Capt. Dan Stauffer](mailto:%20finchasersportfishing@gmail.com)
-   *Fin Chaser* mate
:::
:::

## Schedule

### 15 August

-   Scientific crew (DHS, MOB)
    -   pack CBL truck (2013 Silverado) at Solomons, MD. Drive to West Ocean City, MD
    -   purchase dry ice at Safeway in Chester, MD
        -   1925 Main St, Chester, MD 21619
    -   Check into [Beach Bum West-O](https://beachbumwesto.com/)
        -   12308 Old Bridge Road, Ocean City, Maryland 21842
        -   Conf. #: LAMACMD03207533

### 16 August

-   Scientific crew (DHS, MOB)
    -   Check out of Beach Bum West-O
-   Full crew (DHS, MOB, TS, DS, mate)
    -   0630 - 1500: Recreational fishing survey aboard F/V Fin Chaser
        -   312 Talbot St, Ocean City, MD 21842
-   Scientific crew returns to Solomons, MD. Unpack and return truck.

\newpage

## Cruise plan

Times calculated at 12 knots.

::: columns
::: {.column width="50%"}
-   0500: MOB, DHS, TS meet at Talbot St Pier, load gear onto *F/V Fin Chaser*
-   0530: *Fin Chaser* depart Talbot St Pier, transit to **Great Eastern Reef** control site
    -   <font size= "5"> **38 25.829, 74 46.062** </font>
-   0655: Arrive Great Eastern Reef control site.
    -   Mate arranges gear for jigging.
    -   Sci crew organizes MS-222 bucket
    -   Environmental data recorded
-   0705: Mate begins jigging for 15 minutes.
    -   All fish are weighed and measured.
    -   First 10 black sea bass are sacrificed.
-   0720 Mate ends jigging.
:::

::: {.column width="50%"}
-   0725: Survey crew begins drop survey.
    -   15 x 3 minute drops.
    -   All fish are weighed and measured.
    -   First 10 black sea bass are sacrificed.
-   0900: Drop survey ends. Transit to **K07** experimental site.
    -   <font size= "5"> **38 21.201, 74 42.15** </font>
-   0930: Arrive K07 experimental site.
-   0940: Mate begins jigging for 15 minutes. Data and specimen collection as above.
-   0955: Mate ends jigging.
-   1000: Survey crew begins drop survey, conducted as above.
-   1130: Drop survey ends. Return to Talbot St Pier.
-   1305: Arrive Talbot St Pier. Unload gear.
:::
:::

```{r}
#| echo: false

library(ggplot2); library(sf); library(geoarrow)

bathy <- read_sf('data/geo/bathymetry_midatl.gpkg') |> 
  st_transform(4326)

midatl <- read_geoparquet_sf('data/geo/matl_states.parquet')

ggplot() +
  geom_sf(data = bathy[bathy$elev_m < 100,]) +
  geom_sf(data = midatl) +
  geom_sf(data = stations, size = 6, fill = 'red') +
  geom_sf_label(data = stations, aes(label = station), nudge_x = 0.03, nudge_y = 0.01) +
  geom_sf(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[c(path),])),
      'LINESTRING'
      ), crs = 4326
    )) +
  labs(x = NULL, y = NULL) +
  coord_sf(xlim  = c(-75.2, -74.55),
           ylim = c(38.1, 38.5), expand = F) +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 45))

```

```{r}
#| label: dist-matrix
nm_matrix |>
  as.numeric() |>
  matrix(ncol = 3, dimnames = list(stations$station, stations$station)) |> 
  knitr::kable('html', align = 'c', caption = 'Distance, nautical miles')

time_matrix |>
  matrix(ncol = 3, dimnames = list(stations$station, stations$station)) |> 
  knitr::kable('html', align = 'c', caption = 'Distance, time')
```
