---
title: 'Joint MMM-FRM equipment tending'
format: html
editor: visual
echo: false
message: false
warning: false
params:
  seed:
    value:
      spring: 20240401
      earlysummer: 20240601
      latesummer: 20240801
      autumn: 20241001
  n_control: 2
  n_exp: 4
  vessel_sp_kts: 8
---

```{r}
#| label: params
library(sf); library(ggplot2); library(concaveman); library(TSP); library(units)
library(data.table); library(targets)
```

```{r}
#| label: geo-base

turbines <- usw_locations_202211 |> 
  tar_read() |> 
  st_read(quiet = TRUE)
turbines <- turbines[!grepl('Within 1 nm', turbines$propTSS),]

turbine_boundary <- concaveman::concaveman(turbines)

setDT(turbines)[c(st_contains(turbine_boundary,
                           st_buffer(geom, units::as_units(10, 'meters')),
                           sparse = F)), location := 'inner']
turbines[is.na(location), location := 'outer']


zones <- st_read(tar_read(usw_buildout),
                 quiet = T)
zones <- zones[zones$zone == 'A', ]

ctrl_area <- st_read(tar_read(frm_control),
                     quiet = T)
```

```{r}
#| label: mmm-sites

mmm <- fread(
  text = 
    "station, lat, lon
    T-1C, 38.3319, -74.9491
    A-5C, 38.3360, -74.7226
    D1, 38.2973, -74.6928
    D2, 38.2698, -74.6741
    V4, 38.3350, -74.6768
    V5, 38.3368, -74.7682
    "
) |> 
  st_as_sf(coords = c('lon', 'lat'),
           crs = 4326)
```

```{r}
#| label: site-selection

# create empty lists of control and experimental sites.
# each element will be one cruise
ctrl_sites <- vector("list", 4)
exp_sites <- vector("list", 4)

for(i in seq_along(params$seed)) {
  set.seed(params$seed[[i]])
  
  ctrl_sites[[i]] <- ctrl_area |> 
    st_transform(26918) |>
    st_buffer(dist = -units::set_units(550, 'meter')) |>
    st_sample(params$n_control) |> 
    st_as_sf() |> 
    st_transform(4326) |>
    data.table() |> 
    _[order(st_coordinates(x)[, 1])] |> 
    _[, station := paste0('C', substr(params$seed[[i]], 3, 6),
                          '-',
                          1:params$n_control)] 
  
  setnames(ctrl_sites[[i]], 'x', 'geometry')
  st_geometry(ctrl_sites[[i]]) <- 'geometry'
  
  
  exp_sites[[i]] <- turbines |>
    _[grepl('-1\\d$', name) & location == 'inner'] |>
    _[sample(.N, params$n_exp)] |>
    _[, .(station = name, geometry = geom)] |>
    st_as_sf()
}


```

```{r}
#| label: combine-sites
ocmd <- c(-75.10333, 38.32742) %>% 
  st_point() %>% 
  st_sfc(crs = 4326) %>% 
  st_sf(geometry = .,
        station = 'OCMD')

dm <- function(x){
  paste(as.integer(x), round((abs(x) %% 1) * 60, 4))
}

stations <- vector("list", 4)
for(i in seq_along(params$seed)){
  stations[[i]] <- rbind(ocmd, ctrl_sites[[i]], exp_sites[[i]])
  
  setDT(stations[[i]])[, ':='(lon = st_coordinates(geometry)[,1],
                              lat = st_coordinates(geometry)[,2])]
  
  stations[[i]][, ':='(lon_dm = dm(lon),
                       lat_dm = dm(lat))]
  
  stations[[i]] <- st_as_sf(stations[[i]])
}
```

```{r}
#| label: tsp

tsp <- lapply(stations, 
              function(.) TSP(as.dist(st_distance(.)))
)

tour <- lapply(tsp,
               solve_TSP, method = 'farthest_insertion')

path <- lapply(tour,
               function(.) c(cut_tour(., 1, exclude_cut = FALSE), 1)
)
```

```{r}
#| label: calc-dist-matrices
dist_calc <- function(stations, new_unit = 'nautical_mile'){
  dist_mat <- st_distance(stations)
  
  dist_mat <- set_units(dist_mat, new_unit, mode = 'standard')
  
  rownames(dist_mat) <- stations$station
  colnames(dist_mat) <- stations$station

  dist_mat
}

tf <- function(dist_mat, speed, pretty = T){
  mat <- dist_mat |>
    as.numeric() |>
    matrix(ncol = sqrt(length(dist_mat)))
  
  if(pretty == T){
  mat <- (mat / speed) * 3600
  
  mat <- ceiling(mat / (5 * 60)) * (5 * 60)
  
  mat <- as.POSIXct(mat,
                    tz = 'UTC',
                    origin = '1970-01-01')
  
  mat <- format(mat, '%H h %M m')
  
  mat <- matrix(mat, ncol = sqrt(length(mat)))
  
  diag(mat) <- '-'
  
  }else{
    mat <- ceiling((mat / speed) * 12) / 12
    
    mat <- matrix(mat, ncol = sqrt(length(mat)))
  }
  
  rownames(mat) <- rownames(dist_mat)
  colnames(mat) <- colnames(dist_mat)
  
  mat
}

nm_matrix <- lapply(stations,
                    function(.) round(dist_calc(.), 2)
)
time_matrix_raw <- lapply(nm_matrix, tf, params$vessel_sp_kts, pretty = F)
time_matrix_pretty <- lapply(nm_matrix, tf, params$vessel_sp_kts)
```



The intent of this document is to aid in planning which MMM devices should be tended in the four FRM cruises. The current considerations are:

  - One FRM cruise each will be conducted during the following periods:
    - Spring (March-May); random site selection seed = 20240401
    - Early Summer (May-July); 20240601
    - Late Summer (July-September); 20240801
    - Autumn (September-November); 20241001
  - MMM moorings will be tended on Day 1 of candidate FRM cruises
  - MMM operations must be conducted after all pots are deployment due to limited deck space
  - 1-2 MMM moorings can be tended, largely dependent on distance from final FRM site and sea state
  - MMM mooring priority will be ranked according to:
    - Ease of access from FRM cruises
    - Spatiotemporal presence of target species
    - Battery life of individual instruments
    - Redundancy with other MMM moorings, the Metocean buoy, and the RTWB
  - A5, V4, and V5 have redundancy with each other and Metocean buoy
  - T1C no redundancy (but easiest to get to)
  - D1, D1 have low redundancy and are comparatively hard to get to

## Map

```{r}
library(leaflet)
wms_enc <- 'https://gis.charttools.noaa.gov/arcgis/rest/services/MCS/ENCOnline/MapServer/exts/MaritimeChartService/WMSServer'
leaflet() |> 
  setView(lng = -74.9, lat = 38.25, zoom = 10) |> 
  addWMSTiles(
    wms_enc,
    layers = c(0:3, 6),
    options = WMSTileOptions(format = "image/png", transparent = T)
  ) |> 
  addPolygons(data = zones, color = 'pink') |>
  addPolygons(data = ctrl_area, color = 'gray')|>
  addCircleMarkers(data = mmm, label = ~ station, color = "black") |> 
  addCircleMarkers(data = stations[[1]], label = ~ station, color = 'red',
                   group = "Spring") |> 
  addPolylines(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[[1]][c(path[[1]]),])),
      'LINESTRING'
    ), crs = 4326
  ), color = 'red', group = "Spring") |>
  addCircleMarkers(data = stations[[2]], label = ~ station, color = 'blue',
                   group = "Early Summer") |> 
  addPolylines(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[[2]][c(path[[2]]),])),
      'LINESTRING'
    ), crs = 4326
  ), color = 'blue', group = "Early Summer") |>
  addCircleMarkers(data = stations[[3]], label = ~ station, color = 'green',
                   group = "Late Summer") |> 
  addPolylines(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[[3]][c(path[[3]]),])),
      'LINESTRING'
    ), crs = 4326
  ), color = 'green', group = "Late Summer") |>
  addCircleMarkers(data = stations[[4]], label = ~ station, color = 'yellow',
                   group = "Autumn") |> 
  addPolylines(data = st_sfc(
    st_cast(
      do.call(c, st_geometry(stations[[4]][c(path[[4]]),])),
      'LINESTRING'
    ), crs = 4326
  ), color = 'yellow', group = "Autumn") |>
  addMiniMap(zoomLevelFixed = 7) |> 
  addLayersControl(
    overlayGroups = c("Spring", "Early Summer", "Late Summer", "Autumn"),
    options = layersControlOptions(collapsed = FALSE)
  ) |> 

  addMeasure(primaryLengthUnit = 'meters', primaryAreaUnit = 'hectares')
```


## Summary of map
### Spring

Starting at the farthest-southern site and working northward, FRM deployment will finish at site UH-11.

  - Return transit is directly over A-5C, V5, and T-1C
  - Approximately 20 minutes round-trip (RT) from V4 at 8 knots
  - 21 minutes one-way from D1
  - 37 minutes one-way from D2

### Early Summer

Starting at the farthest-southern site and working northward, FRM deployment will finish at site UH-10.

  - Co-located with A5-C
  - Return transit is directly over V5 and T1-C
  - Approximately 40 minutes round-trip (RT) from V4 at 8 knots
  - 23 minutes one-way from D1
  - 40 minutes one-way from D2
  
### Late Summer

Starting at the farthest-southern site and working northward, FRM deployment will finish at site UK-13.

  - 7 minutes to D1, en route to OCMD
  - 16 minutes from D1 to D2
  
### Autumn

Starting at the farthest-southern site and working northward, FRM deployment will finish at site UL-10.

  - Not very close to any MMM site
  - 32 minutes RT to D1
  - 47 minutes RT to D2

## Overall summary

  - D1 and D2 would be best fit for the Late-Summer (July-September) FRM cruise
    - 3-5 months from initial deployment
  - T1-C can be tended during any FRM cruise
  - A5-C, V4, and V5 would be best fit in the Spring or Early-Summer FRM cruises
